\propertitle

### 2020-01-13 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2020-01-13T15:39:05+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Inferences about large neuronal populations}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The context we consider is as follows. During an experimental session we
have recorded the spiking activities of $n$ neurons for a certain amount of
time. These neurons are our \enquote{sample}. Their spikes are binned into
$T$ time bins, and a binary value 1 is assigned to each bin containing at
least one spike, 0 to the bins without spikes. Call $\ya_{t}$ the number of
neurons that fire during time bin $t$: this is the \emph{total activity} of
the sample, or just \enquote{activity} for short. Obviously
$\ya_{t} \in\set{0,1,\dotsc,n}$; if $\ya_{t}=0$, no neuron spikes during
bin $t$; if $\ya_{t}=n$, all spike (at least once) at some point during bin
$t$, and so on. For brevity we say \enquote{at $t$} for \enquote{during
  time bin
  $t$}. % If we divide the total activity by the population size we have the
% \emph{normalized total activity} or population-averaged activity $\ya/n$,
% ranging from $0$ to $1$ in $1/n$ steps.
From the sequence of activities
$(\ya_{t})\equiv (\ya_{1}, \dotso, \ya_{T})$ we can count how often the
activity levels $\ya=0$, $\ya=1$, and so on appeared during the recording,
obtaining the distribution of measured relative frequencies
$(\yff_{\ya}) \equiv (\yff_{0},\dotso, \yff_{n}) \defs \yf$. We can also
consider the (unknown) sample activity at time bins \emph{outside} of the
recorded period.

@@

The total activity of the larger population of $N$ neurons at $t$ is
$\yA_{t}$. The relative frequencies of the various activity levels during
the recording were
$(\yFF_{\yA}) \equiv (\yFF_{0},\dotso,\yFF_{N}) \defs \yF$. The values
$\yA_{t}$ at each $t$ and the frequency distribution $\yF$ are unknown to
us. We only know for certain that $\yA_{t} \in \set{0, 1, \dotsc, N}$, that
$\yA_{t} \ge \ya_{t}$, and that $N-\yA_{t} \ge n-\ya_{t}$ for obvious
reasons. For the time being we assume that we know $N$; in
\sect~\ref{sec:N} we discuss the consequences of our lack of precise
knowledge about this number.

@@

Our question is whether some amount of \enquote{cooperation} is more
\enquote{important} than another amount or not. In terms of informational
sufficiency, its definition might suggest to compare the belief
distributions conditional on two different subsets of statistics. But which
measure should be used for such a comparison? and how to motivate it? We
prefer to take a different approach, leading to an automatic answer: let us
thus turn the informational sufficiencies of two different subsets of
statistics into two \emph{hypotheses}. We can then ask: which hypothesis is
more probable, given the data? Equivalently, what is the ratio of their
probabilities? By translating our question about \enquote{cooperativity}
into a hypothesis-testing problem about sufficiency, we can find a
well-defined answer determined by the probability-calculus. % The ratio of
% the two probabilities gives us a natural measure of the
% For example, do
% the data say that the first two empirical moments are probably enough to
% determine our beliefs? or is it more probable that the first four are
% needed instead?
We shall see that this approach is indeed more powerful than a simple
comparison of distributions.

@@

The answer comes by examining the origin of our belief distribution for the
activity of the sample \emph{conditional} on the activity of the larger
population. It is given by the hypergeometric
distribution~\eqref{eq:conditional_hypergeometric}, reprinted here:
\begin{equation}
  \labelbis{eq:conditional_hypergeometric}
  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
    \binom{\yA}{\ya}\binom{N-\yA}{n-\ya}\binom{N}{n}^{-1}
\equiv \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}.
\end{equation}
This distribution leads to the equality of factorial
moments~\eqref{eq:equal_m_factorialmoments}, on which our model rests. This
conditional probability, characteristic of \enquote{drawing without
  replacement}
\citep[\eg][\chap~3]{jaynes1994_r2003}, % [\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}
ensues when we believe, with the same degree, that any of the $N$ units
(balls or neurons) could be drawn or sampled. Thus, consider the pool of
all neurons which we believe -- with equal degree of belief for each neuron
-- could have been recorded. Then $N$ is the size of that pool. $N$
therefore depends on factors such as: the shape, dimensions, and technical
specifications of the recording probe; the inaccuracy in the insertion of
the probe, leading for example to slightly different insertion points or
angles; the density of neurons around the probe. But it also depends on the
homogeneity of the brain region where the recording is made: if we believe
that it does not matter whether the probe had been inserted in some other
point of the same brain region, then
formula~\eqref{eq:conditional_hypergeometric} is appropriate. $N$ can
therefore only be assessed case by case.

Whether this symmetry in our beliefs is justified, or for which sampling
procedures it can be justified, is the fundamental, deep question of
sampling theory, which we cannot discuss here \citep[see \eg\ the
discussions and references
in][]{ericson1969,smith1976}. % {kruskaletal1979c,kruskaletal1980}
It is of course possible to probabilistically assess, case by case, whether
this symmetry assumption is appropriate. But we must also be aware that
such assessment in turn rests on analogous symmetry assumptions at a higher
level, and so on. The probability-calculus, just like the truth-calculus,
cannot yield conclusions without premisses
\citep{hailperin2011}[p.~182]{johnson1924}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Acknowledgements
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{acknowledgements}
  This work is financially supported by the Kavli Foundation and the Centre
  of Excellence scheme of the  Research Council of Norway (Yasser Roudi group).\\
  PGLPM thanks the staff of the NTNU library for their always prompt
  support; Mari, Miri, \amp\ Emma for continuous encouragement and
  affection; Buster Keaton and Saitama for filling life with awe and
  inspiration; the developers and maintainers of \LaTeX, Emacs, AUC\TeX,
  Open Science Framework, Python, Inkscape, Sci-Hub for making a free and
  unfiltered scientific exchange possible.
%\rotatebox{15}{P}\rotatebox{5}{I}\rotatebox{-10}{P}\rotatebox{10}{\reflectbox{P}}\rotatebox{-5}{O}.
%\sourceatright{\autanet}
\end{acknowledgements}

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2020-01-13T20:37:41+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Inferences about large neuronal populations}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The context we consider is as follows. During an experimental session we
have recorded the spiking activities of $n$ neurons for a certain amount of
time. These neurons are our \enquote{sample}. Their spikes are binned into
$T$ time bins, and a binary value 1 is assigned to each bin containing at
least one spike, 0 to the bins without spikes. Call $\ya_{t}$ the number of
neurons that fire during time bin $t$: this is the \emph{total activity} of
the sample, or just \enquote{activity} for short. Obviously
$\ya_{t} \in\set{0,1,\dotsc,n}$; if $\ya_{t}=0$, no neuron spikes during
bin $t$; if $\ya_{t}=n$, all spike (at least once) at some point during bin
$t$, and so on. For brevity we say \enquote{at $t$} for \enquote{during
  time bin
  $t$}. % If we divide the total activity by the population size we have the
% \emph{normalized total activity} or population-averaged activity $\ya/n$,
% ranging from $0$ to $1$ in $1/n$ steps.
From the sequence of activities
$(\ya_{t})\equiv (\ya_{1}, \dotso, \ya_{T})$ we can count in how many bins
each activity level $\ya=0$, $\ya=1$, \ldots, $\ya=n$ appeared during the
recording. Dividing each number of bins by the total number of bins $T$ we
obtain the distribution of measured relative frequencies
$(\yff_{0},\dotso, \yff_{n}) \equiv (\yff_{\ya}) \defs \yf$. We can also
consider the (unknown) sample activity at time bins \emph{outside} of the
recorded period.

@@

The total activity of the larger population of $N$ neurons at $t$ is
$\yA_{t}$. The relative frequencies of the various activity levels during
the recording were
$(\yFF_{0},\dotso,\yFF_{N}) \equiv (\yFF_{\yA}) \defs \yF$. The values
$\yA_{t}$ at each $t$ and the frequency distribution $\yF$ are unknown to
us. We only know for certain that $\yA_{t} \in \set{0, 1, \dotsc, N}$, that
$\yA_{t} \ge \ya_{t}$, and that $N-\yA_{t} \ge n-\ya_{t}$ for obvious
reasons. For the time being we assume that we know $N$; in
\sect~\ref{sec:N} we discuss the consequences of our lack of precise
knowledge about this number.

@@

Our question is whether some degree of \enquote{cooperation} is more
\enquote{important} than some other degree, or not. In terms of
informational sufficiency, its definition might suggest to compare the
belief distributions conditional on two different subsets of statistics.
But which measure should be used for such a comparison? and how to motivate
it? We prefer to take a different approach, leading to an automatic answer:
let us thus turn the informational sufficiencies of two different subsets
of statistics into two \emph{hypotheses}. We can then ask: which hypothesis
is more probable, given the data? Equivalently, what is the ratio of their
probabilities? By translating our question about \enquote{cooperativity}
into a hypothesis-testing problem about sufficiency, we can find a
well-defined answer determined by the probability-calculus. % The ratio of
% the two probabilities gives us a natural measure of the
% For example, do
% the data say that the first two empirical moments are probably enough to
% determine our beliefs? or is it more probable that the first four are
% needed instead?
We shall see that this approach is indeed more powerful than a simple
comparison of distributions.

@@

The answer comes by examining the origin of our belief distribution for the
activity of the sample \emph{conditional} on the activity of the larger
population. It is given by the hypergeometric
distribution~\eqref{eq:conditional_hypergeometric}, reprinted here:
\begin{equation}
  \labelbis{eq:conditional_hypergeometric}
  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
    \binom{\yA}{\ya}\binom{N-\yA}{n-\ya}\binom{N}{n}^{-1}
\equiv \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}.
\end{equation}
This distribution leads to the equality of factorial
moments~\eqref{eq:equal_m_factorialmoments}, on which our model rests. This
conditional probability, characteristic of \enquote{drawing without
  replacement}
\citep[\eg][\chap~3]{jaynes1994_r2003}, % [\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}
ensues when we believe, with the same degree, that any of the $N$ units
(balls or neurons) could be drawn or sampled. Thus, consider the pool of
all neurons which we believe -- with an \emph{equal} degree of belief for
each neuron -- could have been recorded. Then $N$ is the size of that pool.
$N$ therefore depends on factors such as: the shape, dimensions, and
technical specifications of the recording probe; the inaccuracy in the
insertion of the probe, leading for example to slightly different insertion
points or angles; the density of neurons around the probe. But it also
depends on the homogeneity of the brain region where the recording is made:
if we believe that it does not matter whether the probe had been inserted
in some other point of the same brain region, then
formula~\eqref{eq:conditional_hypergeometric} is appropriate. $N$ can
therefore only be assessed case by case.

Whether this equality in our degrees of belief is justified, or for which
sampling procedures it can be justified, is the fundamental, deep question
of sampling theory, which we cannot discuss here \citep[see \eg\ the
discussions and references
in][]{ericson1969,smith1976}. % {kruskaletal1979c,kruskaletal1980}
It is of course possible to probabilistically assess, case by case, whether
this symmetry assumption is appropriate. But we must also be aware that
such assessment in turn rests on analogous symmetry assumptions at a higher
level, and so on. The probability-calculus, just like the truth-calculus,
cannot yield conclusions without premisses
\citep{hailperin2011}[p.~182]{johnson1924}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Acknowledgements
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{acknowledgements}
  This work is financially supported by the Kavli Foundation and the Centre
  of Excellence scheme of the  Research Council of Norway (Yasser Roudi group).\\
  We thank Iv\'an Davidovich for very insightful and helpful comments on a
  previous draft of this manuscript. PGLPM thanks the staff of the NTNU
  library for their always prompt support; Mari, Miri, \amp\ Emma for
  continuous encouragement and affection; Buster Keaton and Saitama for
  filling life with awe and inspiration; the developers and maintainers of
  \LaTeX, Emacs, AUC\TeX, Open Science Framework, Python, Inkscape, Sci-Hub
  for making a free and unfiltered scientific exchange possible.
%\rotatebox{15}{P}\rotatebox{5}{I}\rotatebox{-10}{P}\rotatebox{10}{\reflectbox{P}}\rotatebox{-5}{O}.
%\sourceatright{\autanet}
\end{acknowledgements}
